---
title: "Coursera - Regression Models course - Course Project 1"
author: "Martin Cote"
date: "May 24, 2015"
output: pdf_document
fontsize: 7pt
geometry: margin=1.5cm
---

# Executive Summary
Using the "mtcars" dataset (a standard dataset provided with R), we try to answer, using statistical regresssion models, the following questions:

* Is an automatic or manual transmission better for MPG
* Quantify the MPG difference between automatic and manual transmissions

# Pre-Processing

* Loading libraries (GGPLOT2 & CAR required);
* Loading the required data ("mtcars", provided with R); and
* Cleaning the data (i.e. switching 'am' to a factor variable - for specific processing only).

**Note: All pre-processing is hidden.**
```{r echo=FALSE}
library(ggplot2)
library(car)
data(mtcars)

# Understanding the data: (disabled for final report)
#dim(mtcars)
#names(mtcars)
#sapply(mtcars, class)
#head(mtcars, 10)

# Converting 'am' to a factor variable:
mtcars_f <- mtcars # backup
mtcars_f$am <- factor(mtcars_f$am, levels=c(0, 1), labels=c("Automatic", "Manual"))
```

# Exploratory Data Analysis
## Analysis: Is transmission (i.e. 'am', 0 = automatic and 1 = manual) a good predictor for the 'mpg'?
Using first a plot to demonstrate the relation and validate if there are, if any, possible relation between the two random variables (refer to figure 1).

## Inference: Validating the hypothesis using a Student **t-test**
```{r echo=TRUE}
test <- t.test(mpg ~ am, data = mtcars_f, paired=FALSE, var.equal=FALSE)
test$p.value
test$conf.int
```
We observe the p-value of the **Student's Test Distribution** is far below 0.05 which indicates there is a solid relation between Transmission (i.e. 'am') and MPG (i.e. 'mpg'), giving a confidence interval of -11 to -3 MPG at 95% (i.e. a decrease).

# Regression Models
## Linear Model between Transmission and MPG
First, we will validate the linear model between the two desired variables to study.
```{r echo=TRUE}
fit <- lm(mpg ~ am, data = mtcars_f)
summary(fit)$coef # printing out only the coefficients
```
As we can see, there is indeed a linear relationship (which we can be observed in figure 2 as well).  However, to investigate the validity of this linear model, we investigate the residuals (via plot) and confirm that the variability is much greater for the manual transmission (see firgure 3), which could potentially further investigation required (see *selecting the linear model* section).

## Quantifying the differences
Using the linear model (without intercept for quick validation), we can obtain the differences of the impacts on both:
```{r echo=TRUE}
fit.nointer <- lm(mpg ~ am - 1, data = mtcars_f)
summary(fit.nointer)$coef # printing out only the coefficients
```
On average, a manual car will provide a ~7 MPG difference avantage over an automatoic (~24 MPG versus ~17 MPG).

## Diagnostics
Since data entry was performed by hand, we use the handy "leverage measures" to validate if any data is an outlier.  The test confirms that none of the *hatvalues* are too different (i.e. 2, 3 times greater than the mean - '0' means 'FALSE'):
```{r echo=FALSE}
hv <- as.data.frame(hatvalues(fit))
mn <-mean(hatvalues(fit))
hv$warn <- ifelse(hv[, 'hatvalues(fit)'] >= (3*mn), TRUE, # check if greater/equal than 3*
   ifelse(hv[, 'hatvalues(fit)'] >= (2*mn), TRUE, FALSE )) # check if greater/equal than 2*, other FALSE.

sum(hv$warn)
```

## Further investigations: selecting models
Since the residual investigation confirmed that other variables might be at play, it is suggested to look at other models.  Looking at all variables within a linear model, we can see that 'qsec' and 'wt' seems to have an impact (**note**: hidden for quicker read and brief).
```{r echo=FALSE, message=FALSE}
#summary(lm(mpg ~ . -1, data = mtcars))$coef
```
Since these variables have an impact, we then study a second regression model:
```{r echo=TRUE}
summary(lm(mpg ~ am + wt + qsec, data = mtcars))$coef
```
As we can observe from the coefficients, 'wt' (the weight) seems to have a greater impact overall although all of them have a solid relation using the p-values calculated.

\pagebreak

# Appendix

## Figure 1 - Exploratory Analysis (MPG by Transmission)
```{r echo=FALSE, fig.height = 4, fig.width = 4}
# Plot the MPG changes by the type of transmission
ggplot(mtcars_f, aes(x=am, y=mpg)) +
  geom_point(color="black", size=5, alpha=1/2) +
  geom_point(color="red", size=3, alpha=1/2) +
  xlab("Type of Transmission") +
  ylab("MPG") +
  labs(title="Exploratory Analysis (MPG by Transmission)") +
  theme_bw()

```

## Figure 2 - Linear Regression Model for MPG by Transmission
```{r echo=FALSE, fig.height = 4, fig.width = 4}
ggplot(data = mtcars, aes(x = am, y = mpg)) +
  geom_smooth(method = "lm", colour = "black") + # linear model regression
  scale_x_continuous(breaks = c(0, 1),
                     name = "Transmission",
                     labels = c("0" = "Automatic", "1" = "Manual")) +
  geom_point(size = 5, colour = "black", alpha = 0.4) + # black contour
  geom_point(size = 3, colour = "red", alpha = 0.4) + # red center
  ylab("MPG") +
  labs(title="Linear Regression Model (MPG by Transmission") +
  theme_bw()
```

## Figure 3 - Verifying variability within the residual
```{r echo=FALSE, fig.height = 4, fig.width = 4}
ggplot(data = mtcars_f, aes(x = am, y = resid(lm(mpg ~ am))) ) +
  geom_hline(yintercept = 0, size = 2) +
  geom_point(size = 5, colour = "black", alpha = 0.4) + # black contour
  geom_point(size = 3, colour = "red", alpha = 0.4) + # red center
  xlab("Transmission") +
  ylab("Residual") +
  labs(title="Verifying variability within the residuals") +
  theme_bw()
```

## Figure 4 - Diagnostics
```{r echo=FALSE, fig.height=4, fig.width=4}
par(mfrow = c(2, 2))
plot(fit)
```

