---
title: "Coursera - Regression Models course - Course Project 1"
author: "Martin Cote"
date: "May 24, 2015"
output: pdf_document
---

# Executive Summary

Using the "mtcars" dataset provided within the R environment, answering the following questions:

* Is an automatic or manual transmission better for MPG
* Quantify the MPG difference between automatic and manual transmissions

# Pre-Processing

Loading libraries and data.

```{r echo=FALSE}
#library(dplyr)
#library(tidyr)
library(ggplot2)

data(mtcars)
#carsdata <- tbl_df(mtcars)
```

# Exploratory Data Analysis

Is transmission (i.e. 'am', 0 = automatic and 1 = manual) a good predictor for the 'mpg'?  Using a plot to demonstrate the relation and the linear regression between the variables.

```{r echo=TRUE}
# linear regression
lm(mpg ~ am, data=mtcars)

g <- ggplot(mtcars, aes(am, mpg)) +
  geom_point(color = "steelblue", size=4, alpha=1/2) +
  #geom_smooth(size=2, method="lm", formula=mtcars$mpg ~ mtcars$am) +
  xlab("Transmission") +
  ylab("Mile-Per-Gallon") +
  labs(title="Relation between Transmission vs MPG") +
  theme_bw()
g

t.test(mtcars$am, mtcars$mpg, paired=FALSE, var.equal=FALSE)

```

We observe the p-value of the Student's Test Distribution is far below 0.05 which indicates there is a strong relation between Transmission (i.e. 'am') and MPG (i.e. 'mpg').

# Predictions and Correlation

```{r echo=TRUE}
cor(mtcars$mpg, mtcars$am)
var(mtcars$mpg, mtcars$am)

```

